Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-03-25 paper1_jiturra.Rmd
descriptive plots

 # install.packages("trackdown")
 trackdown::upload_file(file = "paper1_jiturra.Rmd", hide_code = TRUE,gpath = '001_PhD/002dissertation/paper1')

@@

```{r}

```{r}

@@

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  religion,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

@@

```{r fig1, fig.dim=c(10,5),fig.cap='Linear Predictions for Network Homogeneity on Redistributive Preferences by Social Class',warning=FALSE}
int_homo <- digclas1_VW_7
df_pred1 <- 
  predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"),
              newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 0.1),
                                 dclass3res_V = levels(dfreg$dclass3res_V)))
# names(df_pred1)
df_pred1 %>% 
ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=2,color="black") +
  facet_grid(~factor(dclass3res_V, levels=
                       c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"),
                     c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))) +
  scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_y_continuous(limits = c(50,80))+
  labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)),
       x="Network homogeneity",y="Redistributive preferences") +
  theme(legend.position = "none", legend.title = element_blank())
```

@@

```{r fig2, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)

>>>>

 # install.packages("trackdown")
 trackdown::update_file(file = "paper1_jiturra.Rmd", hide_code = TRUE,gpath = '001_PhD/002dissertation/paper1')

@@

```{r fig1, fig.dim=c(12,5),fig.cap='Bivariate relationships between Income Inequality, Social Class and Network homogeneity',warning=FALSE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         # gini_disp=scale(gini_disp),
         )

homo_ineq <- dfreg %>%
  # filter(country3 != "ZAF") %>%
  group_by(country2,country3,dclass3res_V) %>%
  summarise_at(c("homclass3_V_res","egal","gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame()

homo_ineq$country <- countrycode::countrycode(sourcevar = homo_ineq$country2,origin = "iso3c",destination = "iso2c")

homo_ineq$dclass3res_V <- forcats::fct_rev(homo_ineq$dclass3res_V)
homo_ineq %>%
ggplot(aes(y=homclass3_V_res, x=gini_disp, group=dclass3res_V,fill=dclass3res_V)) +
  geom_text(label=as.character(homo_ineq$country),size=3) +
  geom_smooth(method = "lm",fullrange=TRUE, color="black",linetype = 2,size=1,se = FALSE) +
  facet_wrap(~dclass3res_V) +
  xlab("Income Inequality")+
  ylab("Network homogeneity") +
  # labs(title = "Class homogeneity and Gini (post-taxes)") +
  scale_fill_manual(values=c("#CCCCCC","#989898","#323232"))+
  scale_color_manual(values=c("#CCCCCC","#989898","#323232")) +
  theme(legend.position = "none",legend.title = element_blank())

@@

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  religion,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

@@

```{r fig3, fig.dim=c(10,5),fig.cap='Linear Predictions for Network Homogeneity on Redistributive Preferences by Social Class',warning=FALSE}
int_homo <- digclas1_VW_7
df_pred1 <- 
  predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"),
              newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 0.1),
                                 dclass3res_V = levels(dfreg$dclass3res_V)))
# names(df_pred1)
df_pred1 %>% 
ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=2,color="black") +
  facet_grid(~factor(dclass3res_V, levels=
                       c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"),
                     c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))) +
  scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_y_continuous(limits = c(50,80))+
  labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)),
       x="Network homogeneity",y="Redistributive preferences") +
  theme(legend.position = "none", legend.title = element_blank())
```

@@

```{r fig4, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)
